---
title: "Lab 3: Displaying Data Visualization on a Website"
author: "Britney Pepper"
output: html_document
date: "2024-09-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(dplyr)
library(tidyverse)
```

</br>

## Introduction

</br>

Installing Packages

```{r install packages}
# install.packages("palmerpenguins")

library("palmerpenguins")

data(package = 'palmerpenguins')
```

</br>

Load and view the data set Palmer Penguins

```{r, load and view the data}
data_raw <- palmerpenguins::penguins_raw
print(data_raw)

data <- palmerpenguins::penguins
print(data)
```

</br>

## Exercise

</br>

#### Data
Loading the data
```{r, loading the data}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
library(dplyr)
library(tidyverse)
library(palmerpenguins)
```

</br>

#### Single trait distribution
Explore the distribution in any one of the morphological traits in the penguin dataset.

```{r, single trait distribution, warning=FALSE, echo=FALSE}
#Density of each bill length of all the data
ggplot(data, aes(x = bill_length_mm)) +
  geom_density()

#Density of each bill length by species
ggplot(data) +
  geom_density(aes(x = bill_length_mm, fill = species))

#Density of each bill length by species to show overlap
ggplot(data) +
  geom_density(aes(x = bill_length_mm, fill = species), alpha = 0.4)
```

We can see by the overlap that the Gentoo and Chinstrap have similar bill length while the Adelie has a small bill length overall.

</br>
</br>

#### Relationship between multiple traits
Explore the relationship between multiple different morphological traits between species in the dataset.

```{r, relationships between multiple traits, warning=FALSE, echo=FALSE, message=FALSE}
#Comparing body mass and bill length
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point()

#Comparing body mass and bill length colored by species
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = species))

#Comparing body mass and bill length colored by species with trend lines
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species))

#Comparing body mass and bill length colored by species with linear trend lines
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species), method = "lm")

#Comparing body mass and flipper length
ggplot(data, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point()

#Comparing body mass and flipper length colored by species
ggplot(data, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species))

#Comparing body mass and flipper length colored by species with trend lines
ggplot(data, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species))

#Comparing body mass and flipper length colored by species with linear trend lines
ggplot(data, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species), method = "lm")
```

</br>

#### Sexual dimorphism
Explore the level of sexual dimorphism in different morphological traits in penguins.

```{r, sexual dimorphism, echo=FALSE, message=FALSE, warning=FALSE}
#Sexual dimorphism for body mass based on sex
ggplot(data, aes(x = sex, y = body_mass_g)) +
  geom_point()
  
#Sexual dimorphism for body mass and bill length based on sex
ggplot(data) +
  geom_point(aes(x = body_mass_g, y = bill_length_mm, color = sex))

#Sexual dimorphism for body mass and bill length based on sex with trend lines
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex)) +
  geom_smooth(aes(color = sex))

#Sexual dimorphism for body mass and bill length based on sex with trend lines
ggplot(data, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex)) +
  geom_smooth(aes(color = sex), method = "lm")

#Comparing by species
ggplot(data, aes(x = bill_length_mm, y=bill_depth_mm, color=sex)) + 
  geom_point() +
  facet_wrap(~species)
```

</br>


